---
output:
  github_document:
    html_preview: true
    fig_width: 8
    fig_height: 8
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE
)
```

# realtalk

<!-- badges: start -->
[![R-CMD-check](https://github.com/wilkox/realtalk/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/wilkox/realtalk/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

`realtalk` is an R package for interacting with the OpenAI Realtime API. It provides functionality for bidirectional streaming of text and audio to and from the OpenAI API.

**Note**: This package is primarily developed for personal use but is being shared in case others find it useful. It is in early development and the API may change.

## Installation

You can install the development version of `realtalk` from GitHub with:

```{r}
# install.packages("devtools")
devtools::install_github("wilkox/realtalk")
```

## Requirements

- A valid [OpenAI API key](https://platform.openai.com/api-keys) with access to the Realtime API
- The [sox](https://sox.sourceforge.net/) command-line utility for audio capture and processing

## Basic Usage

### Creating a Stream

The `Stream` class is the main interface to the OpenAI Realtime API:

```{r}
library(realtalk)

# Create a new streaming session
stream <- Stream$new()

# Start streaming with your API key and preferred model
stream$start_streaming(
  api_key = Sys.getenv("OPENAI_API_KEY"), # Default is openai_api_key()
  model = "gpt-4o-realtime-preview-2025-06-03",
  voice = "ballad" # Choose from available OpenAI TTS voices
)
```

### Sending Text Messages

Send text messages to the API using the `send_text` method:

```{r}
# Send text as a user
stream$send_text("Hello, I'd like to discuss AI safety.", role = "user", trigger_response = TRUE)

# Wait for the complete response
stream$wait_for_response()
```

### Using Audio

The stream automatically captures and sends audio from your microphone, and plays audio responses. For additional audio utility functions:

```{r}
# Capture audio to a file
capture_audio_chunk_file("my_recording.wav", duration = 3)

# Play an audio chunk
audio_b64 <- base64enc::base64encode("my_recording.wav")
play_audio_chunk(audio_b64)
```

### Accessing the Conversation

You can access the complete conversation history at any time:

```{r}
# Get a tibble of all messages in the conversation
conversation <- stream$conversation()

# Print a nicely formatted transcript
stream$transcript()
```

### Ending a Session

When you're done, close the stream:

```{r}
stream$stop_streaming()
```

## Event Logging

The package includes an event logging system that captures all API interactions:

```{r}
# Access event log from stream
events <- stream$events$as_tibble()

# Display events in a readable format
print(events)
```

## Further Documentation

For more detailed information, check the package documentation and vignettes:

```{r}
help(package = "realtalk")
browseVignettes("realtalk")
```

## License

This project is licensed under the MIT License - see the LICENSE file for details.
